---
title: "Read data from web"
author: "Jinghan Zhao"
date: "2024-10-10"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
library(rvest)
library(httr)
```

## Extracting tables

```{r}
url = "https://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url)
```

Get the pieces I actually need.

```{r}
marj_use_df = 
  drug_use_html %>% 
  html_table() %>% 
  first() %>% 
  slice(-1)
```

Read in cost of living data.

```{r}
nyc_cost_df = 
  read_html("https://www.bestplaces.net/cost_of_living/city/new_york/new_york") %>% 
  html_table(header = TRUE) %>% 
  first()
```

## CSS selectors!!

```{r}
swm_url = "https://www.imdb.com/list/ls070150896/"

swm_html = read_html(swm_url)
```

Read html only once, and use the item `swm_html`. (Don't go to a page too many times)

```{r}
swm_title_vec = 
  swm_html %>% 
  html_elements(".ipc-title-link-wrapper .ipc-title__text") %>% 
  html_text()

swm_runtime_vec = 
  swm_html %>% 
  html_elements(".dli-title-metadata-item:nth-child(2)") %>% 
  html_text()

swm_score_vec = 
  swm_html %>% 
  html_elements(".metacritic-score-box") %>% 
  html_text()

swm_df = 
  tibble(
    title = swm_title_vec,
    runtime = swm_runtime_vec,
    score = swm_score_vec
  )
```

Let's import some books.

```{r}
books_html = read_html("https://books.toscrape.com/")

books_html %>% 
  html_elements(".product_pod a") %>% 
  html_text()
```


## Use API

Get water data.

```{r}
nyc_water = 
  GET("https://data.cityofnewyork.us/resource/ia2d-e54m.csv") %>% 
  content()
```

Get BRFSS data.

`$limit` can work for THIS API. Other APIs may have different instructions.

```{r}
brfss_df = 
  GET("https://chronicdata.cdc.gov/resource/acme-vg9e.csv",
      query = list("$limit" = 5000)) %>% 
  content()
```


Pokemon API


More complicated API. 

To build a Pokemon dataset for analysis, you’d need to distill the data returned from the API into a useful format; iterate across all pokemon; and combine the results.

```{r}
pokemon = 
  GET("https://pokeapi.co/api/v2/pokemon/ditto") %>% 
  content()

pokemon$height
pokemon$abilities
```


