---
title: "Strings and Factors"
author: "Jinghan Zhao"
date: "2024-10-15"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Load the necessary libraries.

```{r}
library(rvest)
library(p8105.datasets)
```



## Let's do strings

```{r}
string_vec = c("my", "name", "is", "one")

str_detect(string_vec, "a")
str_detect(string_vec, "One")

str_replace(string_vec, "one", "One")
str_replace(string_vec, "e", "E")
```

Looking for beginning / end of the line.

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")

str_detect(string_vec, "^i think")
str_detect(string_vec, "i think$")
```

Either capital or lowercase.

```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec, "pumpkin")

str_detect(string_vec, "[Pp]umpkin")
```

Looking for a pattern: number followed by letter.

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]")
```

Special: `.` matches everything

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

Looking for protected character: add `\\`

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```



## Factors ...

```{r}
sex_vec = factor(c("male", "male", "female", "female"))

as.numeric(sex_vec)
```


Do some releveling ...

No change in label, but the underlying structure changed.

To see more string or factor related functions: type `str_` or `fct_'

```{r}
sex_vec = fct_relevel(sex_vec, "male")

as.numeric(sex_vec)
```


## NSDUH

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = read_html(url)
```

Get the pieces I actually need.

```{r}
marj_use_df = 
  drug_use_html %>% 
  html_table() %>% 
  first() %>% 
  slice(-1) %>% 
  select(-contains("P Value")) %>% 
  pivot_longer(
    cols = -State,
    names_to = "age_year",
    values_to = "percent"
  ) %>% 
  separate(age_year, into = c("age", "year"), sep = "\\(") %>% 
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_remove(percent, "[a-c]$"),
    percent = as.numeric(percent)
  )
```

`fct_reorder()`: reorder factor levels by sorting along another variable (if one-to multiple, will take the median)

```{r}
marj_use_df %>% 
  filter(age == "12-17",
         !(State %in% c("Total U.S.", "South"))) %>% 
  mutate(
    State = fct_reorder(State, percent)
  ) %>% 
  ggplot(aes(x = State, y = percent, color = year)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


## NYC Restaurant Inspections

```{r}
data("rest_inspec")

rest_inspec %>% 
  slice(1:1000) %>% 
  view()
```

```{r}
rest_inspec %>% 
  count(boro, grade) %>% 
  pivot_wider(
    names_from = grade,
    values_from = n
  )
```

```{r}
rest_inspec = 
  rest_inspec %>% 
  filter(
    str_detect(grade, "[A-C]"),
    !(boro == "Missing")
  )
```

Looking for pizza places

```{r}
rest_inspec %>% 
  mutate(dba = str_to_sentence(dba)) %>% 
  filter(str_detect(dba, "Pizza"))
```

Order in frequency

```{r}
rest_inspec %>% 
  mutate(dba = str_to_sentence(dba)) %>% 
  filter(str_detect(dba, "Pizza")) %>% 
  mutate(boro = fct_infreq(boro)) %>% 
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```

Alphabetical, but pick STATEN ISLAND to the front

```{r}
rest_inspec %>% 
  mutate(dba = str_to_sentence(dba)) %>% 
  filter(str_detect(dba, "Pizza")) %>% 
  mutate(boro = fct_relevel(boro, "STATEN ISLAND")) %>% 
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```

Rename a borough, resulting in wrong order ...

```{r}
rest_inspec %>% 
  mutate(dba = str_to_sentence(dba)) %>% 
  filter(str_detect(dba, "Pizza")) %>% 
  mutate(
    boro = fct_infreq(boro),
    boro = str_replace(boro, "MANHATTAN", "THE CITY")
    ) %>% 
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```

Use `fct_recode()` to fix it

```{r}
rest_inspec %>% 
  mutate(dba = str_to_sentence(dba)) %>% 
  filter(str_detect(dba, "Pizza")) %>% 
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "THE CITY" = "MANHATTAN")
    ) %>% 
  ggplot(aes(x = boro, fill = grade)) +
  geom_bar()
```


One last thing on factors...

Factor will be converted into numbers

```{r}
rest_inspec %>% 
  mutate(dba = str_to_sentence(dba)) %>% 
  filter(str_detect(dba, "Pizza")) %>%
  mutate(boro = fct_infreq(boro)) %>% 
  lm(zipcode ~ boro, data = .)
```

